---
title: "Capstone1"
output: html_document
date: "2026-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## R Markdown

# Download train data
```{r}

df <- read.csv("train.csv")
```

# Calculate Categorical P-values
```{r}
pvals <- sapply(names(df)[sapply(df, function(x) is.factor(x) | is.character(x))], function(x) {
  fit <- aov(as.formula(paste("SalePrice ~", x)), data = df)
  summary(fit)[[1]][["Pr(>F)"]][1]
})

pvals
```

# Determine Categorical significance based on CI 99% 
```{r}
alpha <- 0.01

significance <- ifelse(pvals < alpha, "Significant", "Not Significant")

anova_results <- data.frame(
  Variable = names(pvals),
  P_value = round(pvals, 6),
  Significance = significance
)

anova_results
```
# Identify Not Significant Categorical Variables 
```{r}
not_significant_cat <- anova_results[anova_results$Significance == "Not Significant", ]

print(not_significant_cat)
```



# Compute Numerical P-values 
```{r}
numeric_vars <- names(df)[sapply(df, is.numeric) & names(df) != "SalePrice"]

numeric_pvals <- sapply(numeric_vars, function(var) {
  fit <- lm(as.formula(paste("SalePrice ~", var)), data = df)
  summary(fit)$coefficients[2, "Pr(>|t|)"]
})

numeric_results <- data.frame(
  Variable = names(numeric_pvals),
  P_value = round(numeric_pvals, 6)
)

print(numeric_results)
```

# Determine Numerical significance based on CI 99% 
```{r}
alpha <- 0.01

significance <- ifelse(numeric_results$P_value < alpha, "Significant", "Not Significant")

numeric_results$Significance <- significance

print(numeric_results)
```

# Identify Not Significant Numerical Variables 
```{r}
not_significant_vars <- numeric_results[numeric_results$Significance == "Not Significant", ]

print(not_significant_vars)
```


# Remove Insignificant Variables
```{r}
insignificant_vars <- c(
  "Id", "LandSlope", "Condition2", "BsmtFinType2",
  "PoolQC", "MiscFeature", "Street", "Utilities",
  "BsmtFinSF2", "LowQualFinSF", "BsmtHalfBath",
  "X3SsnPorch", "MiscVal", "MoSold", "YrSold", "BldgType", "HouseStyle"
)

clean.df <- df[, !(names(df) %in% insignificant_vars)]
```


# Determine Reliability Of Significant Categorical Variables
```{r}
cat_vars <- names(clean.df)[sapply(clean.df, function(x) is.factor(x) | is.character(x))]

na_percent <- sapply(clean.df[cat_vars], function(x) {
  sum(is.na(x)) / length(x) * 100
})

na_percent_df <- data.frame(
  Variable = names(na_percent),
  NA_Percent = round(na_percent, 2)
)

na_percent_df <- na_percent_df[order(-na_percent_df$NA_Percent), ]

print(na_percent_df)
```

# Identify Unreliable Categorical Variables (Threshold >30% NA=Unreliable) And Remove
```{r}
vars_to_remove <- names(na_percent[na_percent > 30])
clean.df <- clean.df[, !(names(clean.df) %in% vars_to_remove)]

print(vars_to_remove)
```

# Determine Reliability Of Significant Numerical Variables
```{r}
numeric_vars <- names(clean.df)[sapply(clean.df, is.numeric)]

na_percent_num <- sapply(clean.df[numeric_vars], function(x) {
  sum(is.na(x)) / length(x) * 100
})

numeric_reliability <- data.frame(
  Variable = names(na_percent_num),
  NA_Percent = round(na_percent_num, 2),
  Reliable = ifelse(na_percent_num <= 20, "Yes", "No")  # using 20% threshold
)

numeric_reliability <- numeric_reliability[order(-numeric_reliability$NA_Percent), ]

print(numeric_reliability)
```


# Identify Unreliable Numerical Variables (Threshold >20% NA=Unreliable)
```{r}
vars_to_remove_num <- names(na_percent_num[na_percent_num > 20])

clean.df <- clean.df[, !(names(clean.df) %in% vars_to_remove_num)]

print(vars_to_remove_num)
```

# Remove NA Values Within Categorical Variables Using Mode
```{r}
cat_vars <- names(clean.df)[sapply(clean.df, function(x) is.factor(x) | is.character(x))]

get_mode <- function(x) {
  ux <- na.omit(unique(x))
  ux[which.max(tabulate(match(x, ux)))]
}

for (var in cat_vars) {
  clean.df[[var]][is.na(clean.df[[var]])] <- get_mode(clean.df[[var]])
}
```


# Remove NA Values Within Numerical Variables Using Median
```{r}
numeric_vars <- names(clean.df)[sapply(clean.df, is.numeric)]

for (var in numeric_vars) {
  median_val <- median(clean.df[[var]], na.rm = TRUE)
  clean.df[[var]][is.na(clean.df[[var]])] <- median_val
}
```

```{r}
clean.df
```


```{r}
nprice <- aggregate(SalePrice ~ Neighborhood, data = clean.df, mean)  |>
  arrange(desc(SalePrice))
nprice
```

Identify the Number of Houses Within Each Neighborhood
```{r}
clean.df |>
  count(Neighborhood, sort = TRUE)
```

Separate Neighborhoods into High, Middle, and Low house prices   (Not Used)
```{r}
neighborhood.price <- aggregate(SalePrice ~ Neighborhood, data = clean.df, mean)

neighborhood.price$Price.Level <- cut(
  neighborhood.price$SalePrice,
  breaks = c(98576.47, 177482.75, 256389.03, 335295.32),
  labels = c("Low", "Middle", "High"),
  include.lowest = TRUE
)
neighborhood.price <- neighborhood.price |>
  arrange(desc(SalePrice))

neighborhood.price
```

Convert Neighborhood Names Into Low, Medium, High In Correspondence To Mean House Price (Not Used)
```{r}
clean.df$Neighborhood <- ifelse(clean.df$SalePrice <= 177482.75, "Low",
                         ifelse(clean.df$SalePrice <= 256389.03, "Middle", "High"))
```


```{r}
clean.df
```


Add GarageCars^2 
```{r}
clean.df$GarageCars_sq <- clean.df$GarageCars^2
```


Convert Categorical Variables into Numeric Variables
```{r}
clean.df$ExterQual <- as.numeric(dplyr::recode(clean.df$ExterQual,
  "Po" = 1,
  "Fa" = 2,
  "TA" = 3,
  "Gd" = 4,
  "Ex" = 5
))

#Convert ExterCond to Numeric
clean.df$ExterCond <- as.numeric(dplyr::recode(clean.df$ExterCond,
  "Po" = 1,
  "Fa" = 2,
  "TA" = 3,
  "Gd" = 4,
  "Ex" = 5
))

#Convert BsmtQual to Numeric
clean.df$BsmtQual <- as.numeric(dplyr::recode(clean.df$BsmtQual,
  
  "Po" = 1,
  "Fa" = 2,
  "TA" = 3,
  "Gd" = 4,
  "Ex" = 5
))

#Convert BsmtCond to Numeric
clean.df$BsmtCond <- as.numeric(dplyr::recode(clean.df$BsmtCond,
  "NA" = 0,
  "Po" = 1,
  "Fa" = 2,
  "TA" = 3,
  "Gd" = 4,
  "Ex" = 5
))

#BsmtExposure
clean.df$BsmtExposure <- as.numeric(dplyr::recode(clean.df$BsmtExposure,
  "NA" = 0,
  "No" = 1,
  "Mn" = 2,
  "Av" = 3,
  "Gd" = 4
))

#HeatingQC
clean.df$HeatingQC <- as.numeric(dplyr::recode(clean.df$HeatingQC,
  "Po" = 1,
  "Fa" = 2,
  "TA" = 3,
  "Gd" = 4,
  "Ex" = 5
))

#KitchenQual
clean.df$KitchenQual <- as.numeric(dplyr::recode(clean.df$KitchenQual,
  "Po" = 1,
  "Fa" = 2,
  "TA" = 3,
  "Gd" = 4,
  "Ex" = 5
))

#GarageFinish
clean.df$GarageFinish <- as.numeric(dplyr::recode(as.character(clean.df$GarageFinish),
  "NA" = 0,
  "Unf" = 1,
  "RFn" = 2,
  "Fin" = 3
))

#GarageQual
clean.df$GarageQual <- as.numeric(dplyr::recode(as.character(clean.df$GarageQual),
  "NA" = 0,
  "Po" = 1,
  "Fa" = 2,
  "TA" = 3,
  "Gd" = 4,
  "Ex" = 5
))

#GarageCond
clean.df$GarageCond <- as.numeric(dplyr::recode(as.character(clean.df$GarageCond),
  "NA" = 0,
  "Po" = 1,
  "Fa" = 2,
  "TA" = 3,
  "Gd" = 4,
  "Ex" = 5
))

#BsmtFinType1
clean.df$BsmtFinType1 <- as.numeric(dplyr::recode(as.character(clean.df$BsmtFinType1),
  "NA" = 0,
  "Unf" = 1,
  "LwQ" = 2,
  "Rec" = 3,
  "BLQ" = 4,
  "ALQ" = 5,
  "GLQ" = 6
))

#Convert Functional to Numerical
clean.df$Functional <- as.numeric(dplyr::recode(as.character(clean.df$Functional),
  "Sal" = 0,
  "Sev" = 0,
  "Maj2" = 1,
  "Maj1" = 1,
  "Mod" = 1,
  "Min2" = 2,
  "Min1" = 2,
  "Typ" = 3
))
```


Test for Redundant Variables
```{r}
round(prop.table(table(RoofMatl= clean.df$RoofMatl)) * 100, 2)

round(prop.table(table(LandContour = clean.df$LandContour)) * 100, 2)

round(prop.table(table(Condition1 = clean.df$Condition1)) * 100, 2)

round(prop.table(table(RoofStyle = clean.df$RoofStyle)) * 100, 2)

round(prop.table(table(MasVnrType = clean.df$MasVnrType)) * 100, 2)

round(prop.table(table(ExterCond = clean.df$ExterCond)) * 100, 2)

round(prop.table(table(BsmtCond = clean.df$BsmtCond)) * 100, 2)

round(prop.table(table(BsmtExposure = clean.df$BsmtExposure)) * 100, 2)

round(prop.table(table(Heating = clean.df$Heating)) * 100, 2)

round(prop.table(table(HeatingQC = clean.df$HeatingQC)) * 100, 2)

round(prop.table(table(CentralAir = clean.df$CentralAir)) * 100, 2)

round(prop.table(table(Electrical = clean.df$Electrical)) * 100, 2)

round(prop.table(table(Functional = clean.df$Functional)) * 100, 2)

round(prop.table(table(GarageType = clean.df$GarageType)) * 100, 2)

round(prop.table(table(PavedDrive = clean.df$PavedDrive)) * 100, 2)

round(prop.table(table(SaleType = clean.df$SaleType)) * 100, 2)

round(prop.table(table(Exterior1st = clean.df$Exterior1st)) * 100, 2)

round(prop.table(table(Exterior2nd = clean.df$Exterior2nd)) * 100, 2)

round(prop.table(table(Fireplaces = clean.df$Fireplaces)) * 100, 2)


```

Group rare Variable Categories into "Other" using <3% Threshold  (Not Used)
```{r}
#Exterior1st
freq <- prop.table(table(clean.df$Exterior1st))
rare <- names(freq[freq < 0.03])
clean.df$Exterior1st <- ifelse(clean.df$Exterior1st %in% rare,
                               "Other",
                               clean.df$Exterior1st)

#Exterior2nd
freq <- prop.table(table(clean.df$Exterior2nd))
rare <- names(freq[freq < 0.03])
clean.df$Exterior2nd <- ifelse(clean.df$Exterior2nd %in% rare,
                               "Other",
                               clean.df$Exterior2nd)

#MSZoning
freq <- prop.table(table(clean.df$MSZoning))
rare <- names(freq[freq < 0.03])
clean.df$MSZoning <- ifelse(clean.df$MSZoning %in% rare,
                               "Other",
                               clean.df$MSZoning)


#LotShape
freq <- prop.table(table(clean.df$LotShape))
rare <- names(freq[freq < 0.03])
clean.df$LotShape <- ifelse(clean.df$LotShape %in% rare,
                               "Other",
                               clean.df$LotShape)

#LotConfig
freq <- prop.table(table(clean.df$LotConfig))
rare <- names(freq[freq < 0.03])
 clean.df$LotConfig <- ifelse(clean.df$LotConfig %in% rare,
                               "Other",
                               clean.df$LotConfig)

#RoofStyle
freq <- prop.table(table(clean.df$RoofStyle))
rare <- names(freq[freq < 0.03])
clean.df$RoofStyle <- ifelse(clean.df$RoofStyle %in% rare,
                               "Other",
                               clean.df$RoofStyle)

#RoofStyle
freq <- prop.table(table(clean.df$RoofStyle))
rare <- names(freq[freq < 0.03])
clean.df$RoofStyle <- ifelse(clean.df$RoofStyle %in% rare,
                               "Other",
                               clean.df$RoofStyle)

#MasVnrType
freq <- prop.table(table(clean.df$MasVnrType))
rare <- names(freq[freq < 0.03])
clean.df$MasVnrType <- ifelse(clean.df$MasVnrType %in% rare,
                               "Other",
                               clean.df$MasVnrType)

#Foundation
freq <- prop.table(table(clean.df$Foundation))
rare <- names(freq[freq < 0.03])
clean.df$Foundation <- ifelse(clean.df$Foundation %in% rare,
                               "Other",
                               clean.df$Foundation)


#Foundation
freq <- prop.table(table(clean.df$Foundation))
rare <- names(freq[freq < 0.03])
clean.df$Foundation <- ifelse(clean.df$Foundation %in% rare,
                               "Other",
                               clean.df$Foundation)

#GarageType
freq <- prop.table(table(clean.df$GarageType))
rare <- names(freq[freq < 0.03])
clean.df$GarageType <- ifelse(clean.df$GarageType %in% rare,
                               "Other",
                               clean.df$GarageType)


#SaleCondition
freq <- prop.table(table(clean.df$SaleCondition))
rare <- names(freq[freq < 0.03])
clean.df$SaleCondition <- ifelse(clean.df$SaleCondition %in% rare,
                               "Other",
                               clean.df$SaleCondition)

#SaleCondition
freq <- prop.table(table(clean.df$SaleCondition))
rare <- names(freq[freq < 0.03])
clean.df$SaleCondition <- ifelse(clean.df$SaleCondition %in% rare,
                               "Other",
                               clean.df$SaleCondition)


```


Test for Redundant Variables
```{r}
round(prop.table(table(RoofMatl= clean.df$RoofMatl)) * 100, 2)

round(prop.table(table(LandContour = clean.df$LandContour)) * 100, 2)

round(prop.table(table(Condition1 = clean.df$Condition1)) * 100, 2)

round(prop.table(table(RoofStyle = clean.df$RoofStyle)) * 100, 2)

round(prop.table(table(MasVnrType = clean.df$MasVnrType)) * 100, 2)

round(prop.table(table(ExterCond = clean.df$ExterCond)) * 100, 2)

round(prop.table(table(BsmtCond = clean.df$BsmtCond)) * 100, 2)

round(prop.table(table(BsmtExposure = clean.df$BsmtExposure)) * 100, 2)

round(prop.table(table(Heating = clean.df$Heating)) * 100, 2)

round(prop.table(table(HeatingQC = clean.df$HeatingQC)) * 100, 2)

round(prop.table(table(CentralAir = clean.df$CentralAir)) * 100, 2)

round(prop.table(table(Electrical = clean.df$Electrical)) * 100, 2)

round(prop.table(table(Functional = clean.df$Functional)) * 100, 2)

round(prop.table(table(GarageType = clean.df$GarageType)) * 100, 2)

round(prop.table(table(PavedDrive = clean.df$PavedDrive)) * 100, 2)

round(prop.table(table(SaleType = clean.df$SaleType)) * 100, 2)

round(prop.table(table(Exterior1st = clean.df$Exterior1st)) * 100, 2)

round(prop.table(table(Exterior2nd = clean.df$Exterior2nd)) * 100, 2)

round(prop.table(table(Fireplaces = clean.df$Fireplaces)) * 100, 2)
```


Remove Redundant Variables Within 85% Threshold
```{r}
redundant.vars <- c(
  "BsmtFinSF1", "BsmtUnfSF", "Condition1", "LandContour", "RoofMatl", "Heating", "CentralAir", "Electrical", "PavedDrive", "X1stFlrSF", "X2ndFlrSF", "Exterior2nd"
)

clean.df <- clean.df[, !(names(clean.df) %in% redundant.vars)]
```

```{r}
clean.df
```


Grouping (Not Used later for some models)
```{r}
#Group all Quality Variables
clean.df$TotalQuality <- round((clean.df$ExterQual +
  clean.df$BsmtQual +
  clean.df$KitchenQual +
  clean.df$GarageQual +
  clean.df$OverallQual +
  clean.df$HeatingQC)/6, 2)

#Group all Condition Variables
clean.df$TotalCond <- round((clean.df$OverallCond +
  clean.df$ExterCond +
  clean.df$BsmtCond +
  clean.df$GarageCond)/4, 2)



#TotBath
clean.df$TotBath <- clean.df$BsmtFullBath +
  ((clean.df$HalfBath)/2) +
  clean.df$FullBath 



```


Implement PCA to Weigh Variables And Group
```{r}

area.vars <- clean.df[, c("LotArea", "GarageArea",
                          "GrLivArea", "TotalBsmtSF")]


pca.area <- prcomp(area.vars, scale. = TRUE)


clean.df$TotArea <- pca.area$x[,1]


pca.area$rotation





area.v <- clean.df[, c("WoodDeckSF", "OpenPorchSF",
                          "EnclosedPorch", "ScreenPorch", "PoolArea")]


pca.a <- prcomp(area.v, scale. = TRUE)


clean.df$TotOutAmenSF <- pca.a$x[,1]


pca.a$rotation
```

```{r}
# PCA Visual for Total Area

library(ggplot2)

# Variables used for TotArea PCA
area.vars <- clean.df[, c("TotalBsmtSF", "X1stFlrSF", "X2ndFlrSF", "GrLivArea")]

# Run PCA
area.pca <- prcomp(area.vars, center = TRUE, scale. = TRUE)

# Variance explained
area.var.explained <- area.pca$sdev^2 / sum(area.pca$sdev^2)

area.scree <- data.frame(
  PC = paste0("PC", 1:length(area.var.explained)),
  Variance_Explained = area.var.explained
)

ggplot(area.scree, aes(x = PC, y = Variance_Explained)) +
  geom_col() +
  geom_text(
    aes(label = paste0(round(Variance_Explained * 100, 1), "%")),
    vjust = -0.5
  ) +
  labs(
    title = "PCA Scree Plot for Total Area Variables",
    x = "Principal Component",
    y = "Proportion of Variance Explained"
  ) +
  theme_minimal()
```



```{r}
clean.df
```

create

Remove Grouped Variables
```{r}
grouped.vars <- c(
  "ExterQual", "BsmtQual", "KitchenQual", "GarageQual", "OverallQual", "HeatingQC", "OverallCond", "ExterCond", "BsmtCond", "GarageCond", "WoodDeckSF", "OpenPorchSF", "EnclosedPorch", "ScreenPorch", "PoolArea", "BsmtFullBath", "FullBath", "HalfBath", "LotArea", "GarageArea", "GrLivArea", "TotalBsmtSF"
)

clean.df <- clean.df[, !(names(clean.df) %in% grouped.vars)]
```


```{r}
write.csv(clean.df, "~/Downloads/clean.df.csv", row.names = FALSE)
```

```{r}
hist(clean.df$SalePrice,
     main = "Distribution of Sale Price",
     xlab = "Sale Price",
     col = "lightblue")
```


```{r}
hist(log(clean.df$SalePrice),
     main = "Histogram of Log Sale Price",
     xlab = "Log Sale Price",
     col = "lightblue",
     border = "black")
```

Perform Log Transformation on Data
```{r}
clean.df$LogSalePrice <- log(clean.df$SalePrice)
```

Split Data Into Train and Valid Using 80-20
```{r}
set.seed(123)

train_index <- sample(1:nrow(clean.df), 0.8 * nrow(clean.df))

train.df <- clean.df[train_index, ]
valid.df <- clean.df[-train_index, ]
```


```{r}
train.df <- train.df %>%
  mutate(across(where(is.character), as.factor))

model <- lm(LogSalePrice ~ ., data = train.df)

summary(model)
anova(model)
```



Create Linear Regression Model for Training Set
```{r}
model <- lm(
  LogSalePrice ~ MSZoning + Neighborhood + YearRemodAdd +
    Exterior1st + Foundation +
    BsmtFinType1 + BedroomAbvGr + TotRmsAbvGrd +
    Fireplaces + GarageCars + SaleCondition +
    GarageCars_sq + TotalQuality + TotalCond +
     TotBath + TotArea,
  data = train.df
)

summary(model)
```


```{r}
library(car)

Anova(model, type = 2)
```

```{r}
formula(model)
```



Remove insignifcant variables according to model
```{r}
insig <- c(
  "KitchenAbvGr", "TotOutAmenSF", "GarageYrBlt", "TotBath", "SalePrice", "SaleType"
)

train.df <- train.df[, !(names(train.df) %in% insig)]
```

Training set prediction
```{r}
train.pred.log <- predict(model, newdata = train.df)

train.results <- data.frame(
  Actual_SalePrice = exp(train.df$LogSalePrice),
  Predicted_SalePrice = exp(train.pred.log)
)

train.results
```

```{r}
train.rmse <- sqrt(mean((train.results$Actual_SalePrice - train.results$Predicted_SalePrice)^2))
train.rmse
```


```{r}
remove_vars <- c(
  "KitchenAbvGr", "TotOutAmenSF", "GarageYrBlt", 
  "TotBath", "SalePrice", "SaleType"
)

valid.df <- valid.df[, !(names(valid.df) %in% remove_vars)]
```


```{r}
valid.pred.log <- predict(model, newdata = valid.df)

valid.results <- data.frame(
  Actual_SalePrice = exp(valid.df$LogSalePrice),
  Predicted_SalePrice = exp(valid.pred.log)
)

valid.rmse <- sqrt(mean((valid.results$Actual_SalePrice - valid.results$Predicted_SalePrice)^2))
valid.rmse
```

```{r}
mean(clean.df$SalePrice)
```





Create CSV File for Cleaned Training Data
```{r}
library(readr)
clean <- write_csv(clean.df, "clean.df.csv")

clean.df <- read.csv("~/Downloads/clean.df.csv")
```

```{r}
head(clean.df)
```






Quantile Regression Model
```{r}
install.packages("quantreg")
library(quantreg)
```


```{r}
library(quantreg)
library(dplyr)

train.df <- train.df %>%
  mutate(across(where(is.character), as.factor))

valid.df <- valid.df %>%
  mutate(across(where(is.character), as.factor))

train.qr <- train.df %>%
  select(-SalePrice)

valid.qr <- valid.df %>%
  select(-SalePrice)

qr_formula <- LogSalePrice ~ MSZoning + Neighborhood + YearBuilt + YearRemodAdd +
  LotFrontage + LotShape + LotConfig + Exterior1st + Foundation +
  BsmtExposure + BsmtFinType1 + BedroomAbvGr + TotRmsAbvGrd +
  Functional + Fireplaces + GarageType + GarageFinish +
  GarageCars + GarageCars_sq + SaleCondition +
  TotalQuality + TotalCond + TotBath + TotArea

qr_model_50 <- rq(
  qr_formula,
  data = train.qr,
  tau = 0.50,
  method = "fn"
)

qr_model_50
```

```{r}

quant.pred.log <- predict(qr_model_50, newdata = valid.qr)

quant.pred.price <- exp(quant.pred.log)

actual.valid.price <- exp(valid.qr$LogSalePrice)

quant.valid.rmse <- sqrt(mean((actual.valid.price - quant.pred.price)^2))

quant.valid.rmse
```

```{r}
library(glmnet)
library(dplyr)

train.df <- train.df %>%
  mutate(across(where(is.character), as.factor))

valid.df <- valid.df %>%
  mutate(across(where(is.character), as.factor))

train.reg <- train.df %>%
  select(-SalePrice)

valid.reg <- valid.df %>%
  select(-SalePrice)

reg_formula <- LogSalePrice ~ MSZoning + Neighborhood +
  YearBuilt + YearRemodAdd + HouseAge + RemodAge +
  LotArea + LotFrontage + LotShape + LotConfig +
  Exterior1st + Exterior2nd + MasVnrArea +
  Foundation + BsmtQual + BsmtCond + BsmtExposure + BsmtFinType1 +
  TotalBsmtSF + X1stFlrSF + X2ndFlrSF + GrLivArea + TotalSF +
  FullBath + HalfBath + BsmtFullBath + BsmtHalfBath + TotalBathrooms +
  BedroomAbvGr + KitchenQual + TotRmsAbvGrd + Functional +
  Fireplaces + FireplaceQu + GarageType + GarageFinish +
  GarageCars + GarageArea + GarageAge +
  HasGarage + HasBasement + HasFireplace +
  WoodDeckSF + OpenPorchSF + EnclosedPorch + ScreenPorch +
  SaleCondition

train.reg$dataset <- "train"
valid.reg$dataset <- "valid"

combined.reg <- rbind(train.reg, valid.reg)

x.combined <- model.matrix(reg_formula, data = combined.reg)[, -1]

x.train <- x.combined[combined.reg$dataset == "train", ]
x.valid <- x.combined[combined.reg$dataset == "valid", ]

y.train <- combined.reg$LogSalePrice[combined.reg$dataset == "train"]
y.valid <- combined.reg$LogSalePrice[combined.reg$dataset == "valid"]
```

```{r}
set.seed(123)

lasso.cv <- cv.glmnet(
  x = x.train,
  y = y.train,
  alpha = 1,
  nfolds = 10
)

lasso.best.lambda <- lasso.cv$lambda.min

lasso.model <- glmnet(
  x = x.train,
  y = y.train,
  alpha = 1,
  lambda = lasso.best.lambda
)

lasso.pred.log <- predict(
  lasso.model,
  s = lasso.best.lambda,
  newx = x.valid
)

lasso.pred.price <- exp(lasso.pred.log)
actual.valid.price <- exp(y.valid)

lasso.valid.rmse <- sqrt(mean((actual.valid.price - lasso.pred.price)^2))
lasso.valid.rmse
```

```{r}
set.seed(123)

ridge.cv <- cv.glmnet(
  x = x.train,
 
   y = y.train,
  alpha = 0,
  nfolds = 10
)

ridge.best.lambda <- ridge.cv$lambda.min

ridge.model <- glmnet(
  x = x.train,
  y = y.train,
  alpha = 0,
  lambda = ridge.best.lambda
)

ridge.pred.log <- predict(
  ridge.model,
  s = ridge.best.lambda,
  newx = x.valid
)

ridge.pred.price <- exp(ridge.pred.log)

ridge.valid.rmse <- sqrt(mean((actual.valid.price - ridge.pred.price)^2))
ridge.valid.rmse
```


```{r}

rmse.comparison <- data.frame(
  Model = c(
    "Linear Regression",
    "Lasso Regression",
    "Ridge Regression",
    "Quantile Regression"
  ),
  Validation_RMSE = c(
    valid.rmse,
    lasso.valid.rmse,
    ridge.valid.rmse,
    quant.valid.rmse
  )
)

rmse.comparison <- rmse.comparison[order(rmse.comparison$Validation_RMSE), ]

rmse.comparison
```

```{r}

library(dplyr)
library(glmnet)
library(quantreg)

vars.to.readd <- c(
  "OverallQual", "OverallCond", "LotArea", "MasVnrArea",
  "BsmtQual", "BsmtCond", "TotalBsmtSF", "YrSold",
  "X1stFlrSF", "X2ndFlrSF", "GrLivArea",
  "FullBath", "HalfBath", "BsmtFullBath", "BsmtHalfBath",
  "KitchenQual", "GarageArea", "GarageYrBlt",
  "WoodDeckSF", "OpenPorchSF", "EnclosedPorch", "ScreenPorch",
  "Exterior2nd", "FireplaceQu"
)
for (v in vars.to.readd) {
  if (!(v %in% names(clean.df)) && v %in% names(df)) {
    clean.df[[v]] <- df[[v]]
  }
}

char.vars <- names(clean.df)[sapply(clean.df, is.character)]

for (v in char.vars) {
  clean.df[[v]][is.na(clean.df[[v]])] <- "None"
}

num.vars <- names(clean.df)[sapply(clean.df, is.numeric)]

for (v in num.vars) {
  clean.df[[v]][is.na(clean.df[[v]])] <- median(clean.df[[v]], na.rm = TRUE)
}


clean.df <- clean.df %>%
  mutate(
    TotalSF = TotalBsmtSF + X1stFlrSF + X2ndFlrSF,
    
    TotalBathrooms = FullBath +
      0.5 * HalfBath +
      BsmtFullBath +
      0.5 * BsmtHalfBath,
    
    HouseAge = YrSold - YearBuilt,
    RemodAge = YrSold - YearRemodAdd,
    GarageAge = YrSold - GarageYrBlt,
    
    HasGarage = ifelse(GarageArea > 0, 1, 0),
    HasBasement = ifelse(TotalBsmtSF > 0, 1, 0),
    HasFireplace = ifelse(Fireplaces > 0, 1, 0),
    
    GarageCars_sq = GarageCars^2,
    LogSalePrice = log(SalePrice)
  )

clean.df$HouseAge[clean.df$HouseAge < 0] <- 0
clean.df$RemodAge[clean.df$RemodAge < 0] <- 0
clean.df$GarageAge[clean.df$GarageAge < 0] <- 0

clean.df <- clean.df %>%
  mutate(across(where(is.character), as.factor))

set.seed(123)

train.index <- sample(
  1:nrow(clean.df),
  size = 0.80 * nrow(clean.df)
)

train.df <- clean.df[train.index, ]
valid.df <- clean.df[-train.index, ]


model_formula <- LogSalePrice ~ MSZoning + Neighborhood +
  OverallQual + OverallCond +
  YearBuilt + YearRemodAdd + HouseAge + RemodAge +
  LotArea + LotFrontage + LotShape + LotConfig +
  Exterior1st + MasVnrArea +
  Foundation + BsmtQual + BsmtCond + BsmtExposure + BsmtFinType1 +
  TotalBsmtSF + X1stFlrSF + X2ndFlrSF + GrLivArea + TotalSF +
  FullBath + HalfBath + BsmtFullBath + BsmtHalfBath + TotalBathrooms +
  BedroomAbvGr + KitchenQual + TotRmsAbvGrd + Functional +
  Fireplaces + FireplaceQu + GarageType + GarageFinish +
  GarageCars + GarageCars_sq + GarageArea + GarageAge +
  HasGarage + HasBasement + HasFireplace +
  WoodDeckSF + OpenPorchSF + EnclosedPorch + ScreenPorch +
  SaleCondition

lm.model <- lm(
  model_formula,
  data = train.df
)

lm.pred.log <- predict(lm.model, newdata = valid.df)

lm.pred.price <- exp(lm.pred.log)

actual.valid.price <- valid.df$SalePrice

lm.valid.rmse <- sqrt(mean((actual.valid.price - lm.pred.price)^2))

train.reg <- train.df %>%
  select(-SalePrice)

valid.reg <- valid.df %>%
  select(-SalePrice)

train.reg$dataset <- "train"
valid.reg$dataset <- "valid"

combined.reg <- rbind(train.reg, valid.reg)

x.combined <- model.matrix(model_formula, data = combined.reg)[, -1]

x.train <- x.combined[combined.reg$dataset == "train", ]
x.valid <- x.combined[combined.reg$dataset == "valid", ]

y.train <- combined.reg$LogSalePrice[combined.reg$dataset == "train"]
y.valid <- combined.reg$LogSalePrice[combined.reg$dataset == "valid"]

actual.valid.price <- exp(y.valid)

set.seed(123)

lasso.cv <- cv.glmnet(
  x = x.train,
  y = y.train,
  alpha = 1,
  nfolds = 10
)

lasso.pred.log <- predict(
  lasso.cv,
  s = "lambda.min",
  newx = x.valid
)

lasso.pred.price <- exp(lasso.pred.log)

lasso.valid.rmse <- sqrt(mean((actual.valid.price - lasso.pred.price)^2))

set.seed(123)

ridge.cv <- cv.glmnet(
  x = x.train,
  y = y.train,
  alpha = 0,
  nfolds = 10
)

ridge.pred.log <- predict(
  ridge.cv,
  s = "lambda.min",
  newx = x.valid
)

ridge.pred.price <- exp(ridge.pred.log)

ridge.valid.rmse <- sqrt(mean((actual.valid.price - ridge.pred.price)^2))


train.qr <- train.df %>%
  select(-SalePrice)

valid.qr <- valid.df %>%
  select(-SalePrice)

quant.model <- rq(
  model_formula,
  data = train.qr,
  tau = 0.50,
  method = "fn"
)

quant.pred.log <- predict(
  quant.model,
  newdata = valid.qr
)

quant.pred.price <- exp(quant.pred.log)

quant.valid.rmse <- sqrt(mean((actual.valid.price - quant.pred.price)^2))

rmse.comparison <- data.frame(
  Model = c(
    "Linear Regression",
    "Lasso Regression",
    "Ridge Regression",
    "Quantile Regression"
  ),
  Validation_RMSE = c(
    lm.valid.rmse,
    lasso.valid.rmse,
    ridge.valid.rmse,
    quant.valid.rmse
  )
)

rmse.comparison <- rmse.comparison[order(rmse.comparison$Validation_RMSE), ]

rmse.comparison

best.model <- rmse.comparison[1, ]

cat(
  "The best model is",
  best.model$Model,
  "with a validation RMSE of",
  round(best.model$Validation_RMSE, 2)
)
```

```{r}
quant.pred.log <- predict(
  quant.model,
  newdata = valid.qr
)

quant.pred.log <- pmin(quant.pred.log, log(max(train.df$SalePrice) * 3))
quant.pred.log <- pmax(quant.pred.log, log(min(train.df$SalePrice) * 0.5))

quant.pred.price <- exp(quant.pred.log)

actual.valid.price <- valid.df$SalePrice

quant.valid.rmse <- sqrt(mean((actual.valid.price - quant.pred.price)^2))

quant.valid.rmse
```

```{r}
rmse.comparison <- data.frame(
  Model = c(
    "Linear Regression",
    "Lasso Regression",
    "Ridge Regression",
    "Quantile Regression"
  ),
  Validation_RMSE = c(
    lm.valid.rmse,
    lasso.valid.rmse,
    ridge.valid.rmse,
    quant.valid.rmse
  )
)

rmse.comparison <- rmse.comparison[order(rmse.comparison$Validation_RMSE), ]

rmse.comparison
```

```{r}
quant_formula <- LogSalePrice ~ Neighborhood +
  OverallQual + OverallCond +
  GrLivArea + TotalSF +
  TotalBathrooms +
  GarageCars + GarageArea +
  YearBuilt + YearRemodAdd +
  HouseAge + RemodAge +
  TotalBsmtSF +
  Fireplaces +
  LotArea

train.qr <- train.df %>%
  select(-SalePrice)

valid.qr <- valid.df %>%
  select(-SalePrice)

quant.model <- rq(
  quant_formula,
  data = train.qr,
  tau = 0.50,
  method = "fn"
)

quant.pred.log <- predict(
  quant.model,
  newdata = valid.qr
)

quant.pred.price <- exp(quant.pred.log)

actual.valid.price <- valid.df$SalePrice

quant.valid.rmse <- sqrt(mean((actual.valid.price - quant.pred.price)^2))

quant.valid.rmse
```

```{r}
rmse.comparison <- data.frame(
  Model = c(
    "Linear Regression",
    "Lasso Regression",
    "Ridge Regression",
    "Quantile Regression"
  ),
  Validation_RMSE = c(
    lm.valid.rmse,
    lasso.valid.rmse,
    ridge.valid.rmse,
    quant.valid.rmse
  )
)

rmse.comparison <- rmse.comparison[order(rmse.comparison$Validation_RMSE), ]

rmse.comparison
```
```{r}
clean.df
```

```{r}
library(dplyr)
library(glmnet)
library(quantreg)

vars.to.readd <- c(
  "OverallQual", "OverallCond", "LotArea", "LotFrontage",
  "MSZoning", "Neighborhood", "LotShape", "LotConfig",
  "Exterior1st", "Foundation", "BsmtExposure", "BsmtFinType1",
  "BedroomAbvGr", "TotRmsAbvGrd", "Functional", "Fireplaces",
  "GarageType", "GarageFinish", "GarageCars", "GarageArea",
  "SaleCondition", "YearBuilt", "YearRemodAdd", "YrSold",
  "TotalBsmtSF", "X1stFlrSF", "X2ndFlrSF", "GrLivArea",
  "WoodDeckSF", "OpenPorchSF", "EnclosedPorch", "ScreenPorch",
  "FullBath", "HalfBath", "BsmtFullBath", "BsmtHalfBath",
  "KitchenQual", "BsmtQual"
)

for (v in vars.to.readd) {
  if (!(v %in% names(clean.df)) && v %in% names(df)) {
    clean.df[[v]] <- df[[v]]
  }
}

char.vars <- names(clean.df)[sapply(clean.df, is.character)]

for (v in char.vars) {
  clean.df[[v]][is.na(clean.df[[v]])] <- "None"
}

num.vars <- names(clean.df)[sapply(clean.df, is.numeric)]

for (v in num.vars) {
  clean.df[[v]][is.na(clean.df[[v]])] <- median(clean.df[[v]], na.rm = TRUE)
}

clean.df <- clean.df %>%
  mutate(
    TotalBathrooms = FullBath +
      0.5 * HalfBath +
      BsmtFullBath +
      0.5 * BsmtHalfBath,
    
    HouseAge = YrSold - YearBuilt,
    RemodAge = YrSold - YearRemodAdd,
    
    HasGarage = ifelse(GarageArea > 0, 1, 0),
    HasBasement = ifelse(TotalBsmtSF > 0, 1, 0),
     HasFireplace = ifelse(Fireplaces > 0, 1, 0),
    
    GarageCars_sq = GarageCars^2,
    LogSalePrice = log(SalePrice)
  )

clean.df$HouseAge[clean.df$HouseAge < 0] <- 0
clean.df$RemodAge[clean.df$RemodAge < 0] <- 0

clean.df <- clean.df %>%
  mutate(across(where(is.character), as.factor))

set.seed(123)

train.index <- sample(
  1:nrow(clean.df),
  size = 0.80 * nrow(clean.df)
)

train.df <- clean.df[train.index, ]
valid.df <- clean.df[-train.index, ]

area.vars <- c(
  "TotalBsmtSF",
  "X1stFlrSF",
  "X2ndFlrSF",
  "GrLivArea"
)

area.pca <- prcomp(
  train.df[, area.vars],
  center = TRUE,
  scale. = TRUE
)

train.area.pca <- predict(area.pca, newdata = train.df[, area.vars])
valid.area.pca <- predict(area.pca, newdata = valid.df[, area.vars])

train.df$TotArea_PC1 <- train.area.pca[, 1]
train.df$TotArea_PC2 <- train.area.pca[, 2]

valid.df$TotArea_PC1 <- valid.area.pca[, 1]
valid.df$TotArea_PC2 <- valid.area.pca[, 2]

outdoor.vars <- c(
  "WoodDeckSF",
  "OpenPorchSF",
  "EnclosedPorch",
  "ScreenPorch"
)

outdoor.pca <- prcomp(
  train.df[, outdoor.vars],
  center = TRUE,
  scale. = TRUE
)

train.outdoor.pca <- predict(outdoor.pca, newdata = train.df[, outdoor.vars])
valid.outdoor.pca <- predict(outdoor.pca, newdata = valid.df[, outdoor.vars])

train.df$TotOutAmen_PC1 <- train.outdoor.pca[, 1]
train.df$TotOutAmen_PC2 <- train.outdoor.pca[, 2]

valid.df$TotOutAmen_PC1 <- valid.outdoor.pca[, 1]
valid.df$TotOutAmen_PC2 <- valid.outdoor.pca[, 2]

pca_formula <- LogSalePrice ~ MSZoning + Neighborhood +
  OverallQual + OverallCond +
  YearBuilt + YearRemodAdd + HouseAge + RemodAge +
  LotArea + LotFrontage + LotShape + LotConfig +
  Exterior1st + Foundation +
  BsmtQual + BsmtExposure + BsmtFinType1 +
  BedroomAbvGr + KitchenQual + TotRmsAbvGrd + Functional +
  Fireplaces + GarageType + GarageFinish +
  GarageCars + GarageCars_sq + GarageArea +
  TotalBathrooms + HasGarage + HasBasement + HasFireplace +
  TotArea_PC1 + TotArea_PC2 +
  TotOutAmen_PC1 + TotOutAmen_PC2 +
  SaleCondition

train.model <- train.df %>%
  select(-SalePrice)

valid.model <- valid.df %>%
  select(-SalePrice)

train.model$dataset <- "train"
valid.model$dataset <- "valid"

combined.model <- rbind(train.model, valid.model)

x.combined <- model.matrix(pca_formula, data = combined.model)[, -1]

x.train <- x.combined[combined.model$dataset == "train", ]
x.valid <- x.combined[combined.model$dataset == "valid", ]

y.train <- combined.model$LogSalePrice[combined.model$dataset == "train"]
y.valid <- combined.model$LogSalePrice[combined.model$dataset == "valid"]

actual.valid.price <- exp(y.valid)

x.train.df <- as.data.frame(x.train)
x.valid.df <- as.data.frame(x.valid)

linear.pca.model <- lm(
  y.train ~ .,
  data = x.train.df
)

linear.pca.pred.log <- predict(
  linear.pca.model,
  newdata = x.valid.df
)

linear.pca.pred.price <- exp(linear.pca.pred.log)

linear.pca.rmse <- sqrt(mean((actual.valid.price - linear.pca.pred.price)^2))

set.seed(123)

lasso.pca.cv <- cv.glmnet(
  x = x.train,
  y = y.train,
  alpha = 1,
  nfolds = 10
)

lasso.pca.pred.log <- predict(
  lasso.pca.cv,
  s = "lambda.min",
  newx = x.valid
)

lasso.pca.pred.price <- exp(lasso.pca.pred.log)

lasso.pca.rmse <- sqrt(mean((actual.valid.price - lasso.pca.pred.price)^2))

set.seed(123)

ridge.pca.cv <- cv.glmnet(
  x = x.train,
  y = y.train,
  alpha = 0,
  nfolds = 10
)

ridge.pca.pred.log <- predict(
  ridge.pca.cv,
  s = "lambda.min",
  newx = x.valid
)

ridge.pca.pred.price <- exp(ridge.pca.pred.log)

ridge.pca.rmse <- sqrt(mean((actual.valid.price - ridge.pca.pred.price)^2))

quant.pca.model <- rq(
  y.train ~ .,
  data = x.train.df,
  tau = 0.50,
  method = "fn"
)

quant.pca.pred.log <- predict(
  quant.pca.model,
  newdata = x.valid.df
)

quant.pca.pred.price <- exp(quant.pca.pred.log)

quant.pca.rmse <- sqrt(mean((actual.valid.price - quant.pca.pred.price)^2))

pca.rmse.comparison <- data.frame(
  Model = c(
    "Linear Regression with PCA",
    "Lasso Regression with PCA",
    "Ridge Regression with PCA",
    "Quantile Regression with PCA"
  ),
  Validation_RMSE = c(
    linear.pca.rmse,
    lasso.pca.rmse,
    ridge.pca.rmse,
    quant.pca.rmse
  )
)

pca.rmse.comparison <- pca.rmse.comparison[
  order(pca.rmse.comparison$Validation_RMSE),
]

pca.rmse.comparison

best.pca.model <- pca.rmse.comparison[1, ]

cat(
  "The best PCA-based model is",
  best.pca.model$Model,
  "with a validation RMSE of",
  round(best.pca.model$Validation_RMSE, 2)
)

quant.pca.model
```

```{r}
calc_r2 <- function(actual, predicted) {
  ss_res <- sum((actual - predicted)^2)
  ss_tot <- sum((actual - mean(actual))^2)
  r2 <- 1 - (ss_res / ss_tot)
  return(r2)
}

linear.pca.r2 <- calc_r2(actual.valid.price, linear.pca.pred.price)
lasso.pca.r2 <- calc_r2(actual.valid.price, lasso.pca.pred.price)
ridge.pca.r2 <- calc_r2(actual.valid.price, ridge.pca.pred.price)
quant.pca.r2 <- calc_r2(actual.valid.price, quant.pca.pred.price)

pca.r2.comparison <- data.frame(
  Model = c(
    "Linear Regression with PCA",
    "Lasso Regression with PCA",
    "Ridge Regression with PCA",
    "Quantile Regression with PCA"
  ),
  Validation_R2 = c(
    linear.pca.r2,
    lasso.pca.r2,
    ridge.pca.r2,
    quant.pca.r2
  )
)

pca.r2.comparison <- pca.r2.comparison[
  order(pca.r2.comparison$Validation_R2, decreasing = TRUE),
]

pca.r2.comparison
```


```{r}

library(quantreg)
library(ggplot2)
library(dplyr)

clean.df$GarageCars_sq <- clean.df$GarageCars^2
clean.df$LogSalePrice <- log(clean.df$SalePrice)

garage.quant.model <- rq(
  LogSalePrice ~ GarageCars + GarageCars_sq,
  data = clean.df,
  tau = 0.50
)

summary(garage.quant.model)

garage.pred.df <- data.frame(
  GarageCars = seq(
    min(clean.df$GarageCars, na.rm = TRUE),
    max(clean.df$GarageCars, na.rm = TRUE),
    length.out = 100
  )
)

garage.pred.df$GarageCars_sq <- garage.pred.df$GarageCars^2

garage.pred.df$Predicted_LogSalePrice <- predict(
  garage.quant.model,
  newdata = garage.pred.df
)

garage.pred.df$Predicted_SalePrice <- exp(garage.pred.df$Predicted_LogSalePrice)

coef.garage <- coef(garage.quant.model)

turning.point <- -coef.garage["GarageCars"] / (2 * coef.garage["GarageCars_sq"])

turning.point
```

```{r}

ggplot(clean.df, aes(x = GarageCars, y = SalePrice)) +
  geom_jitter(alpha = 0.35, width = 0.15, height = 0) +
  geom_line(
    data = garage.pred.df,
    aes(x = GarageCars, y = Predicted_SalePrice),
    linewidth = 1.2
  ) +
  geom_vline(
    xintercept = turning.point,
    linetype = "dashed",
    linewidth = 1
  ) +
  labs(
    title = "Effect of Garage Cars on Sale Price",
    subtitle = paste(
      "Quantile regression with GarageCars and GarageCars squared. Turning point:",
      round(turning.point, 2),
      "garage spaces"
    ),
    x = "Garage Cars",
    y = "Sale Price"
  ) +
  theme_minimal()
```

```{r}
library(ggplot2)
library(dplyr)
library(quantreg)
library(scales)

clean.df$GarageCars_sq <- clean.df$GarageCars^2
clean.df$LogSalePrice <- log(clean.df$SalePrice)

garage.quant.model <- rq(
  LogSalePrice ~ GarageCars + GarageCars_sq,
  data = clean.df,
  tau = 0.50
)

garage.pred.df <- data.frame(
  GarageCars = seq(
    min(clean.df$GarageCars, na.rm = TRUE),
    max(clean.df$GarageCars, na.rm = TRUE),
    length.out = 200
  )
)

garage.pred.df$GarageCars_sq <- garage.pred.df$GarageCars^2

garage.pred.df$Predicted_LogSalePrice <- predict(
  garage.quant.model,
  newdata = garage.pred.df
)

garage.pred.df$Predicted_SalePrice <- exp(garage.pred.df$Predicted_LogSalePrice)

coef.garage <- coef(garage.quant.model)

turning.point <- -coef.garage["GarageCars"] / 
  (2 * coef.garage["GarageCars_sq"])

turning.price <- garage.pred.df$Predicted_SalePrice[
  which.min(abs(garage.pred.df$GarageCars - turning.point))
]

garage.summary <- clean.df %>%
  group_by(GarageCars) %>%
  summarise(
    MedianSalePrice = median(SalePrice, na.rm = TRUE),
    Count = n(),
    .groups = "drop"
  )

ggplot() +
  geom_point(
    data = garage.summary,
    aes(x = GarageCars, y = MedianSalePrice, size = Count),
    alpha = 0.75
  ) +
  geom_line(
    data = garage.pred.df,
    aes(x = GarageCars, y = Predicted_SalePrice),
    linewidth = 1.4
  ) +
  geom_vline(
    xintercept = turning.point,
    linetype = "dashed",
    linewidth = 1
  ) +
  annotate(
    "label",
    x = turning.point,
    y = turning.price,
    label = paste0("Turning point: ", round(turning.point, 2), " cars"),
    vjust = -1,
    size = 4
  ) +
  scale_y_continuous(labels = dollar_format()) +
  scale_x_continuous(
    breaks = sort(unique(clean.df$GarageCars))
  ) +
  scale_size_continuous(name = "Number of Homes") +
  labs(
    title = "Garage Capacity and Predicted Sale Price",
    subtitle = "Median quantile regression with GarageCars and GarageCars²",
    x = "Garage Capacity",
    y = "Sale Price"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

```{r}
library(ggplot2)
library(dplyr)
library(scales)

clean.df$GarageCars_sq <- clean.df$GarageCars^2

garage.summary <- clean.df %>%
  group_by(GarageCars) %>%
  summarise(
    MedianSalePrice = median(SalePrice, na.rm = TRUE),
    MeanSalePrice = mean(SalePrice, na.rm = TRUE),
    Count = n(),
    .groups = "drop"
  )

garage.summary
```

```{r}
library(quantreg)

garage.quant.model <- rq(
  log(SalePrice) ~ GarageCars + GarageCars_sq,
  data = clean.df,
  tau = 0.50
)

garage.pred.df <- garage.summary %>%
  select(GarageCars) %>%
  mutate(GarageCars_sq = GarageCars^2)

garage.pred.df$Predicted_LogSalePrice <- predict(
  garage.quant.model,
  newdata = garage.pred.df
)

garage.pred.df$Predicted_SalePrice <- exp(garage.pred.df$Predicted_LogSalePrice)

garage.plot.df <- garage.summary %>%
  left_join(garage.pred.df, by = "GarageCars")

ggplot(garage.plot.df, aes(x = GarageCars)) +
  geom_line(aes(y = MedianSalePrice), linewidth = 1.4) +
  geom_point(aes(y = MedianSalePrice, size = Count), alpha = 0.85) +
  geom_line(
    aes(y = Predicted_SalePrice),
    linewidth = 1.2,
    linetype = "dashed"
  ) +
  geom_text(
    aes(y = MedianSalePrice, label = dollar(round(MedianSalePrice, 0))),
    vjust = -1,
    size = 4
  ) +
  scale_y_continuous(labels = dollar_format()) +
  scale_x_continuous(breaks = garage.plot.df$GarageCars) +
  scale_size_continuous(name = "Number of Homes") +
  labs(
    title = "Garage Capacity: Actual Median vs Quantile Prediction",
    subtitle = "Solid line = actual median sale price; dashed line = quantile regression prediction",
    x = "Garage Capacity",
    y = "Sale Price"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

```{r}
library(ggplot2)
library(dplyr)
library(scales)

garage.summary <- clean.df %>%
  group_by(GarageCars) %>%
  summarise(
    MedianSalePrice = median(SalePrice, na.rm = TRUE),
    Count = n(),
    .groups = "drop"
  )

ggplot(garage.summary, aes(x = GarageCars, y = MedianSalePrice)) +
  geom_line(linewidth = 1.4) +
  geom_point(aes(size = Count), alpha = 0.85) +
  geom_text(
    aes(label = dollar(round(MedianSalePrice, 0))),
    vjust = -1,
    size = 4
  ) +
  scale_y_continuous(labels = dollar_format()) +
  scale_x_continuous(
    breaks = garage.summary$GarageCars
  ) +
  scale_size_continuous(name = "Number of Homes") +
  labs(
    title = "Median Sale Price by Garage Capacity",
    subtitle = "Sale price increases up to 3 garage spaces, then decreases for 4-car garages",
    x = "Garage Capacity",
    y = "Median Sale Price"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

```{r}
library(dplyr)
library(ggplot2)

linear.summary <- summary(linear.pca.model)

linear.significant <- as.data.frame(linear.summary$coefficients)

linear.significant$Variable <- rownames(linear.significant)

linear.significant <- linear.significant %>%
  rename(
    Estimate = Estimate,
    Std_Error = `Std. Error`,
    T_Value = `t value`,
    P_Value = `Pr(>|t|)`
  ) %>%
  filter(Variable != "(Intercept)") %>%
  mutate(
    Significant = ifelse(P_Value < 0.05, "Yes", "No"),
    Importance = abs(Estimate)
  ) %>%
  arrange(P_Value)

linear.significant.only <- linear.significant %>%
  filter(Significant == "Yes")

linear.significant.only
```

```{r}
log_rmse <- function(actual_price, predicted_price) {
  sqrt(mean((log(actual_price) - log(predicted_price))^2))
}
```

```{r}
linear.pca.log.rmse <- log_rmse(, linear.pca.pred.price)

lasso.pca.log.rmse <- log_rmse(actual.valid.price, lasso.pca.pred.price)

ridge.pca.log.rmse <- log_rmse(actual.valid.price, ridge.pca.pred.price)

quant.pca.log.rmse <- log_rmse(actual.valid.price, quant.pca.pred.price)
```

